---
title: "Assignment 5: Substance Abuse Treatment Mapping"
subtitle: "STUDENT NAME"
author: "Amber Camp"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Tuesday, 10/7 11:59pm**

In this assignment, you will apply analytic and mapping skills on the TEDS-D dataset from SAMHSA.

Refer to the following files for reference, though some of what is asked of you in this assignment will require logic beyond what was discussed here:

-   week06-treatment-mapping-part1.qmd

-   week06-treatment-mapping-part2.qmd

Total time estimate: Two hours

## Load Packages and Data

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(janitor) 
library(arrow) # lets us read/write parquet and feather files
library(sf) # "simple features" for spatial data
library(urbnmapr) # access to US state/country shapefiles for mapping
library(naniar) # tools for exploring and visualizing missing data
library(ggiraph) # makes ggplot charts interactive
options(scipen = 99) # turns off scientific notation

```

```{r}
teds_d <- read_parquet(here("data/teds_d_clean.parquet"))
```

## Question 1

Make an interactive map with `ggiraph` showing the percentage of completed treatments that end with no use at discharge. I am looking for a single interactive map as the final product.

Hint: Break this question down into bite-sized steps, like we did in Part 2. The `freq1_d` column contains information about use at discharge. Check your denominator carefully — should it be *all cases* or *completed cases*?

```{r}

#filtering for only completed
completed <- teds_d %>% 
  filter(reason == "Treatment completed")

#calc % of no use @ discharge by state
percent_completed_by_state <- completed %>%
  group_by(stfips) %>%
  summarize(completed_cases = n(),
    no_use = sum(freq1_d == "no use", na.rm = TRUE)) %>%
  mutate(percentage_no_use = (no_use / completed_cases) * 100,
         state_fips = sprintf("%02d", stfips))

#get mapping data
states_map <- get_urbn_map(map = "states", sf = TRUE)

#joining data
percent_completed_by_state_map <- full_join(states_map, 
                                            percent_completed_by_state,
                                            by = "state_fips")

#plotting map + tooltip
gg_map <- ggplot(percent_completed_by_state_map) +
  geom_sf_interactive(
    aes(
      geometry = geometry,
      fill = percentage_no_use,
      tooltip = paste0(
    state_name, "<br>",
    #"Completed cases: ", completed_cases, "<br>", #to show number of completed cases
    #"No use at discharge: ", no_use, "<br>", #to show number of no use @ discharge
    "Percent no use: ", round(percentage_no_use, 1), "%"
  ),
      data_id = state_fips
    ),
    color = "#ffffff", size = 0.25
  ) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    title = "Completed Treatments Ending with No Use at Discharge",
    fill = "% No Use"
  ) +
  coord_sf(datum = NA) +
  theme_minimal()

#girafe
girafe(
  ggobj = gg_map,
  options = list(
    opts_hover(css = "fill-opacity:0.5;cursor:pointer;"),
    opts_tooltip(css = "background:white;color:black;padding:5px;border-radius:5px;")
  )
)
```

## Question 2

Treatment outcomes can vary depending on the type of service patients receive and the length of stay in treatment.

-   Investigate how the percentage of treatments completed and the percentage of treatments that end with no use at discharge differ when you look at service type and length of stay.

-   Create at least three different visualizations to help answer this question.

-   At least one visualization should focus on service type and at least one should focus on length of stay. A third visualization should compare the two.

-   After making your visualizations, write a short summary (3-ish sentences) that explains the main patterns you observe. Please use full, grammatical sentences.

```{r}

```

Summary: (write here)

## Push to GitHub to Submit!

Assignment is due Tuesday 10/7 at 11:59 pm.
