---
title: "Assignment 4: Exploratory Data Analysis"
author: "Amber Camp"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Sunday, 9/28 11:59pm**

In this assignment, you will practice Exploratory Data Analysis (EDA) by making plots from a **Kaggle dataset of your choice**.

You will submit **at least 4 different plot types**, with short written interpretations for each.

Part 1 is worth a total of 1 point, and Part 2 is worth 4 points (1 point for each plot & interpretation).

## Part 1: Get Your Data

1.  Go to [Kaggle Datasets](https://www.kaggle.com/datasets), create an account, and choose a dataset that interests you. Ensure your data has:
    1.  At least 4 variables
    2.  At least one categorical and one continuous variable
    3.  I recommend data that contains at least 50-100 observations
    4.  Tip: Open the CSV on Kaggle first and skim the columns to check that it meets these requirements
2.  Download the dataset as a CSV file.
3.  Place the CSV in your class R project data folder.
4.  Read it into R with `read_csv()`.

```{r, message=FALSE}
library(tidyverse)

vgdata <- read_csv("data/vg_sales.csv")

```

### Describe your data

Data author: Rush Kirubi

Link: <https://www.kaggle.com/datasets/rush4ratio/video-game-sales-with-ratings?resource=download>

Description: Video game sales from Vgchartz and corresponding ratings from Metacritic

### Preview your data

Use some of the strategies you've learned in this class or picked up elsewhere to summarize and quickly explore your data.

```{r}

skim_data <- skimr::skim(vgdata)
view(skim_data)
```

### Summary stats

Before you make plots, you must also:

-   Show basic summaries of your dataset, including:

    -   `count()` for categorical variables (to see category frequencies)

    -   `mean()`, `median()`, `sd()`, `min()`, `max()` for at least one numeric variable

```{r}

count <- vgdata %>%
  filter(!is.na(Genre), !is.na(Platform)) %>%
  count(Genre, Platform)

count$n

vgdata %>%
  summarise( mean_sales = mean(Global_Sales, na.rm = TRUE),
             median_sales = median(Global_Sales, na.rm = TRUE))
```

Write 2–3 sentences here: What do the category counts tell you? What does the numeric summary show about spread, center, or outliers?

The category counts show which genres, platforms, or publishers are most prominent. The numeric summary essentially shows that the sales are skewed with most games selling very little and some outliers selling far more.

## Part 2:

Create **at least 4 plots**, each of a different type. At least one of your four plots must use different variables than the others.

Each plot should have:

-   A title and labeled axes

-   A short interpretation (2-3 sentences) underneath

Use the EDA Part 3 notes as your reference. Since this is an EDA assignment, your plots can be plain, but consider adding colors, themes, or interactivity for interpretability and visual appeal if you know how.

### Plot 1

```{r}

#point plot
ggplot(vgdata, aes(x = Genre, y = Global_Sales)) +
  geom_count(alpha = 0.2, color = "steelblue") +
   theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Global Sales by Genre (Point Plot)",
       x = "Genre",
       y = "Global Sales (millions)")
```

Interpretation:

Overall the distribution of games released per genre is seemingly even; however, action and adventure games seem to be released more frequently.

### Plot 2

```{r}

#density plot
ggplot(vgdata, aes(x = Global_Sales)) +
  geom_density(fill = "pink", alpha = 0.5) +
  scale_x_log10() + #to better display data bc some games sell exponentially more than others.
  labs(title = "Global Sales",
       x = "Global Sales (millions, log scale)",
       y = "Density") +
  theme_minimal()

```

Interpretation:

This plot shows that a majority of games sold modestly under 1 million units; however, there are some best-selling titles that skew the data as they account for a majority of global sales.

### Plot 3

```{r}

#bar plot
ggplot(vgdata, aes(x = Year_of_Release, fill = Genre)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Number of Games Released per Year by Genre",
       x = "Year of Release",
       y = "Count")
```

Interpretation:

There was a high influx of games being released in the late 2000s and early 2010s with a steep drop off in 2012. Around this time there were many new emerging consoles such as the 3ds or WiiU so it seemingly reflects that in the graph.

### Plot 4

```{r}

#histogram
ggplot(vgdata, aes(x = Critic_Score)) +
  geom_histogram(binwidth = 5, fill = "orange", color = "black") +
  labs(title = "Histogram of Critic Scores",
       x = "Critic Score",
       y = "Count") +
  theme_minimal()
```

Interpretation:

Overall, critics scores vary but there are very little games that achieve a relatively low score or a perfect score. This means that generally the games that are produced are of decent quality considering the histogram is mainly concentrated around games achieving scores around 50-90.

## Push to GitHub to Submit!

Assignment is due Sunday 9/28 at 11:59 pm.
