---
title: "Assignment 9: Class Wrap-up"
format: html
editor: visual
author: "Amber Camp"
date: "`r Sys.Date()`"
---

## Assignment 9: Class Wrap-up

We are finally at the end of the semester! The final assignment for the class will be to summarize and share some of the things we've done so far.

The plan is for everyone to be randomly paired with a tiny topic pertaining to something we've learned together this semester and then submit a code snippet to be presented in class on Wednesday, 11/12. (Think \~2 minute walkthrough)

#### Libraries

```{r, message=FALSE}
library(tidyverse)
```

#### Read in the list of topics

```{r, message=FALSE}
# topics <- read_csv("data/cs201_topics.csv")
# view(topics)

# anything to edit?
# topics[7, "topic"] <- "wildcard"
```

#### Read in the list of student names

```{r, message=FALSE}
# names <- read_csv("data/cs201_names.csv")
# view(names)
```

#### Generate a random number in the names dataframe

```{r, message=FALSE}
# names <- names %>%
#   mutate(number = sample(1:nrow(topics), nrow(names), replace = FALSE))
# 
# view(names)
```

#### Join the names to the topics

```{r, message=FALSE}
# assigned_topics <- names %>%
#   full_join(topics, by = "number")
# 
# print(assigned_topics)
# 
# write.csv(assigned_topics, "data/assignment9_assigned_topics.csv")

# note that if you run the above script and randomize the names again, your outcome will be different from mine!!
# go with the topics assigned in class and announced on Canvas!
```

### Instructions

1.  Find your assigned topic. Talk to me if you reeeeeeaaaally want to do something else, or trade with someone (but let me know).

2.  Prepare a code snippet that briefly (\~2) explains the topic in the code chunk below, using `penguins` data.

3.  Push to GitHub **BY 10:00am** on Wednesday 11/10 (I'll need time to compile before class).

4.  Come to class Wednesday and show us how it's done :)

#### Extra libraries needed (if any)

```{r, message=FALSE}
library(janitor)
library(plotly)
library(ggplot2)
```

#### Read in penguins data

```{r, message=FALSE}
penguins <- read_csv("data/penguins_lter.csv")

penguins <- penguins %>% 
  clean_names()
```

#### facet_wrap() vs. facet_grid()

facet_wrap() allows you to break a plot into multiple smaller plots based on one categorical variable and is arranged into multiple rows and columns. The facets are wrapped and/or rearranged into new rows when they don't fit.

```{r, message=FALSE}
# facet_wrap()

wrap <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(size=0.75, alpha = 0.75) +
  facet_wrap(~ species, # organizes plots based on species
             labeller = labeller(species = c(
               "Adelie Penguin (Pygoscelis adeliae)" = "Adelie",
               "Chinstrap penguin (Pygoscelis antarctica)" = "Chinstrap",
               "Gentoo penguin (Pygoscelis papua)" = "Gentoo"
             ))) +
  labs(
    title = "Penguins: Flipper Length vs Body Mass",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

ggplotly(wrap)
```

facet_grid() allows you to compare ***two*** categorical factors of your data across various rows and columns in a fixed grid. Each combination will get its own panel and it will notwrap or rearrange.

```{r}
# facet_grid()

grid <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(size = 0.75, alpha = 0.5) +
  facet_grid(sex ~ species, # organizes plots in a row-column structure based on sex and species
             labeller = labeller(species = c(
               "Adelie Penguin (Pygoscelis adeliae)" = "Adelie",
               "Chinstrap penguin (Pygoscelis antarctica)" = "Chinstrap",
               "Gentoo penguin (Pygoscelis papua)" = "Gentoo"
             ))) +
  labs(
    title = "Penguins: Flipper Length vs Body Mass",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

ggplotly(grid)
```
